---
title: "Mar653_Final"
author: "Michelle Mak"
date: "12/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Data}
original_data <- read.csv("DMEFExtractSummaryV01.csv")
#sample <- read.csv("modified_dataset.csv")
```

```{r Remove NAs}
str(original_data[,142:159])

# Clean NA's
myDataClean <- na.omit(original_data)
dim(myDataClean)
summary(myDataClean)
colSums(is.na(myDataClean)) #no na's! 
str(myDataClean[,142:159])

#write.csv(myDataClean, file = "Clean_Cust_Data.csv")
```

```{r More PreProcessing}

# Create copy of cleaned data for manipulation
cust_infoDF <-myDataClean

# Convert characters to integers
cust_infoDF$FirstChannel <- as.integer(cust_infoDF$FirstChannel)
cust_infoDF$Email <- as.integer(cust_infoDF$Email)


# drop extra levels and convert to factors
cust_infoDF$Travel[cust_infoDF$Travel==' ']=NA
cust_infoDF$Travel= droplevels(cust_infoDF$Travel)
cust_infoDF$Travel <- as.integer(cust_infoDF$Travel)
str(cust_infoDF$Travel)

cust_infoDF$CurrAff[cust_infoDF$CurrAff==' ']=NA
cust_infoDF$CurrAff= droplevels(cust_infoDF$CurrAff)
cust_infoDF$CurrAff <- as.integer(cust_infoDF$CurrAff)
str(cust_infoDF$CurrAff)

cust_infoDF$CurrEv[cust_infoDF$CurrEv==' ']=NA
cust_infoDF$CurrEv= droplevels(cust_infoDF$CurrEv)
cust_infoDF$CurrEv <- as.integer(cust_infoDF$CurrEv)
str(cust_infoDF$CurrEv)

cust_infoDF$Wines[cust_infoDF$Wines==' ']=NA
cust_infoDF$Wines= droplevels(cust_infoDF$Wines)
cust_infoDF$Wines <- as.integer(cust_infoDF$Wines)
str(cust_infoDF$Wines)

cust_infoDF$FineArts[cust_infoDF$FineArts==' ']=NA
cust_infoDF$FineArts= droplevels(cust_infoDF$FineArts)
cust_infoDF$FineArts <- as.integer(cust_infoDF$FineArts)
str(cust_infoDF$FineArts)

cust_infoDF$Exercise[cust_infoDF$Exercise==' ']=NA
cust_infoDF$Exercise= droplevels(cust_infoDF$Exercise)
cust_infoDF$Exercise <- as.integer(cust_infoDF$Exercise)
str(cust_infoDF$Exercise)

cust_infoDF$SelfHelp[cust_infoDF$SelfHelp==' ']=NA
cust_infoDF$SelfHelp= droplevels(cust_infoDF$SelfHelp)
cust_infoDF$SelfHelp <- as.integer(cust_infoDF$SelfHelp)
str(cust_infoDF$SelfHelp)

cust_infoDF$Collect[cust_infoDF$Collect==' ']=NA
cust_infoDF$Collect= droplevels(cust_infoDF$Collect)
cust_infoDF$Collect <- as.integer(cust_infoDF$Collect)
str(cust_infoDF$Collect)

cust_infoDF$Needle[cust_infoDF$Needle==' ']=NA
cust_infoDF$Needle= droplevels(cust_infoDF$Needle)
cust_infoDF$Needle <- as.integer(cust_infoDF$Needle)
str(cust_infoDF$Needle)

cust_infoDF$Sewing[cust_infoDF$Sewing==' ']=NA
cust_infoDF$Sewing= droplevels(cust_infoDF$Sewing)
cust_infoDF$Sewing <- as.integer(cust_infoDF$Sewing)
str(cust_infoDF$Sewing)

cust_infoDF$DogOwner[cust_infoDF$DogOwner==' ']=NA
cust_infoDF$DogOwner= droplevels(cust_infoDF$DogOwner)
cust_infoDF$DogOwner <- as.integer(cust_infoDF$DogOwner)
str(cust_infoDF$DogOwner)

cust_infoDF$CarOwner[cust_infoDF$CarOwner==' ']=NA
cust_infoDF$CarOwner= droplevels(cust_infoDF$CarOwner)
cust_infoDF$CarOwner <- as.integer(cust_infoDF$CarOwner)
str(cust_infoDF$CarOwner)

cust_infoDF$Cooking[cust_infoDF$Cooking==' ']=NA
cust_infoDF$Cooking= droplevels(cust_infoDF$Cooking)
cust_infoDF$Cooking <- as.integer(cust_infoDF$Cooking)
str(cust_infoDF$Cooking)

cust_infoDF$Pets[cust_infoDF$Pets==' ']=NA
cust_infoDF$Pets= droplevels(cust_infoDF$Pets)
cust_infoDF$Pets <- as.integer(cust_infoDF$Pets)
str(cust_infoDF$Pets)

cust_infoDF$Fashion[cust_infoDF$Fashion==' ']=NA
cust_infoDF$Fashion= droplevels(cust_infoDF$Fashion)
cust_infoDF$Fashion <- as.integer(cust_infoDF$Fashion)
str(cust_infoDF$Fashion)

cust_infoDF$Hunting[cust_infoDF$Hunting==' ']=NA
cust_infoDF$Hunting= droplevels(cust_infoDF$Hunting)
cust_infoDF$Hunting <- as.integer(cust_infoDF$Hunting)
str(cust_infoDF$Hunting)

cust_infoDF$Camping[cust_infoDF$Camping==' ']=NA
cust_infoDF$Camping= droplevels(cust_infoDF$Camping)
cust_infoDF$Camping <- as.integer(cust_infoDF$Camping)

cust_infoDF$Boating[cust_infoDF$Boating==' ']=NA
cust_infoDF$Boating= droplevels(cust_infoDF$Boating)
cust_infoDF$Boating <- as.integer(cust_infoDF$Boating)

cust_infoDF$Child0_2[cust_infoDF$Child0_2==' ']=NA
cust_infoDF$Child0_2= droplevels(cust_infoDF$Child0_2)
cust_infoDF$Child0_2 <- as.integer(cust_infoDF$Child0_2)

cust_infoDF$Child3_5[cust_infoDF$Child3_5==' ']=NA
cust_infoDF$Child3_5= droplevels(cust_infoDF$Child3_5)
cust_infoDF$Child3_5 <- as.integer(cust_infoDF$Child3_5)

cust_infoDF$Child6_11[cust_infoDF$Child6_11==' ']=NA
cust_infoDF$Child6_11= droplevels(cust_infoDF$Child6_11)
cust_infoDF$Child6_11 <- as.integer(cust_infoDF$Child6_11)

cust_infoDF$Child12_16[cust_infoDF$Child12_16==' ']=NA
cust_infoDF$Child12_16= droplevels(cust_infoDF$Child12_16)
cust_infoDF$Child12_16 <- as.integer(cust_infoDF$Child12_16)

cust_infoDF$Child17_18[cust_infoDF$Child17_18==' ']=NA
cust_infoDF$Child17_18= droplevels(cust_infoDF$Child17_18)
cust_infoDF$Child17_18 <- as.integer(cust_infoDF$Child17_18)


str(cust_infoDF[,163:167])
cust_infoDF <- na.omit(cust_infoDF)
colSums(is.na(cust_infoDF))
#cust_infoDF
```

```{r Exploratory Analysis}
# What are the average prices spent on items in the spring vs. the fall? 

## create new dataframe
Fall_Numbers <- cust_infoDF[c("RetF07Dollars", "RetF07Lines", "RetF06Dollars", "RetF06Lines", "RetF05Dollars", "RetF05Lines", "RetF04Dollars", "RetF04Lines", "IntF07GDollars", "IntF07NGDollars", "IntF07Lines", "IntF06GDollars", "IntF06NGDollars", "IntF06Lines", "IntF05GDollars", "IntF05NGDollars", "IntF05Lines", "IntF04GDollars", "IntF04NGDollars", "IntF04Lines", "CatF07GDollars", "CatF07NGDollars", "CatF07Lines", "CatF06GDollars", "CatF06NGDollars","CatF06Lines", "IntF05GDollars", "IntF05NGDollars", "CatF05Lines", "CatF04GDollars", "CatF04NGDollars", "CatF04Lines")] 

Spring_Numbers <- cust_infoDF[c("RetS07Dollars", "RetS07Lines", "RetS06Dollars", "RetS06Lines", "RetS05Dollars", "RetS05Lines", "RetS04Dollars", "RetS04Lines", "IntS07GDollars", "IntS07NGDollars", "IntS07Lines", "IntS06GDollars", "IntS06NGDollars", "IntS06Lines", "IntS05GDollars", "IntS05NGDollars", "IntS05Lines", "IntS04GDollars", "IntS04NGDollars", "IntS04Lines", "CatS07GDollars", "CatS07NGDollars", "CatS07Lines", "CatS06GDollars", "CatS06NGDollars","CatS06Lines", "IntS05GDollars", "IntS05NGDollars", "CatS05Lines", "CatS04GDollars", "CatS04NGDollars", "CatS04Lines")]

## Sum of all columns
colSums(Fall_Numbers)
colSums(Spring_Numbers)


## Avg prices by total order
is.na(Fall_Numbers) <- Fall_Numbers==0 #make 0's na's so they don't negatively skew the avg
colMeans(Fall_Numbers,  na.rm = TRUE, dims = 1)

is.na(Spring_Numbers) <- Spring_Numbers==0
colMeans(Spring_Numbers, na.rm = TRUE, dims = 1)

```

# Average Fall Numbers    
   RetF07Dollars       RetF07Lines     RetF06Dollars       RetF06Lines     RetF05Dollars 
        90.069930          7.599650         66.557434          5.127114         64.106752 
      RetF05Lines     RetF04Dollars       RetF04Lines    IntF07GDollars   IntF07NGDollars 
         5.109968         64.176297          4.810731         98.625974         94.575000 
      IntF07Lines    IntF06GDollars   IntF06NGDollars       IntF06Lines    IntF05GDollars 
         2.599089         81.686688         81.679688          2.035471         81.125201 
  IntF05NGDollars       IntF05Lines    IntF04GDollars   IntF04NGDollars       IntF04Lines 
        73.818182          2.163073         82.072527         72.840796          2.065487 
   CatF07GDollars   CatF07NGDollars       CatF07Lines    CatF06GDollars   CatF06NGDollars 
       109.149002        103.125850          2.658038         95.896410         93.272093 
      CatF06Lines  IntF05GDollars.1 IntF05NGDollars.1       CatF05Lines    CatF04GDollars 
         2.341850         81.125201         73.818182          2.217133         83.419534 
  CatF04NGDollars       CatF04Lines 
        84.558195          2.133705
        
# Average Spring Numbers
    RetS07Dollars       RetS07Lines     RetS06Dollars       RetS06Lines     RetS05Dollars 
        56.543750          5.151042         51.126706          4.661793         48.453029 
      RetS05Lines     RetS04Dollars       RetS04Lines    IntS07GDollars   IntS07NGDollars 
         4.296752         51.701280          4.551920         52.448864         75.885714 
      IntS07Lines    IntS06GDollars   IntS06NGDollars       IntS06Lines    IntS05GDollars 
         1.500000         52.934884         60.800000          1.386282         51.443850 
  IntS05NGDollars       IntS05Lines    IntS04GDollars   IntS04NGDollars       IntS04Lines 
        52.257143          1.482759         48.733871         48.971429          1.428571 
   CatS07GDollars   CatS07NGDollars       CatS07Lines    CatS06GDollars   CatS06NGDollars 
        65.520930         66.278481          1.541353         63.169421         69.328000 
      CatS06Lines  IntS05GDollars.1 IntS05NGDollars.1       CatS05Lines    CatS04GDollars 
         1.510769         51.443850         52.257143          1.582237         49.720812 
  CatS04NGDollars       CatS04Lines 
        78.750000          1.517510 
       


```{r Retail, Internet, Catalog}
retail <- cust_infoDF[3:29]
colSums(retail)

internet <- cust_infoDF[31:66]
colSums(internet)

catalog <- cust_infoDF[68:103]
colSums(catalog)
```
  
```{r K means PreProcessing}
#install.packages("cluster")
library(cluster)
#install.packages("factoextra")
library(factoextra)
#install.packages("mclust")
library(mclust)
#install.packages("tm")
library(tm)
library(knitr)
library(caret)

kmeansdata <- cust_infoDF[,c(136:170)]
kmeansdata$AcqDate = substr(kmeansdata$AcqDate ,1,nchar(kmeansdata$AcqDate )-2) #change acqdate to year of acquisition 
kmeansdata$AcqDate <- as.integer(kmeansdata$AcqDate)

str(kmeansdata)
colSums(is.na(kmeansdata))

scaled_data <- scale(kmeansdata)
head(scaled_data)

#Add Cust_ID back to data
scaled_data  <- cbind(cust_infoDF$Cust_ID, scaled_data)
colnames(scaled_data)[1] <- "Cust_ID"
head(scaled_data)

```

```{r}
# CLUSTERING WITH K-MEANS
k <- kmeans(scaled_data[,c(-1)],2,nstart = 50,iter.max = 15) 

# Clustering with 2 center (Fall vs. Spring Shoppers?)
k2 <- kmeans(scaled_data[,c(-1)],2,nstart = 50,iter.max = 15) # keep number of iter.max=15 to ensure the algorithm converges and nstart=50 to #ensure that at least 50 random sets are choosen  
k2

fviz_cluster(k2, data=scaled_data[,c(-1)],ellipse.type = "convex")+ theme_minimal() # the clusters are hard to depict since it is in 2D. Use a confusion matrix decipher the k-means

```

```{r}
# Elbow Method for finding the optimal number of clusters
set.seed(123)
# Compute and plot wss for k = 2 to k = 8.
k.max <- 8
data <- scaled_data
wss <- sapply(1:k.max, 
              function(k){kmeans(data, k, nstart=50,iter.max = 15 )$tot.withinss})
wss
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")

# looks like the most obvious elbow is at 2! we will create 3 customer segments. 
```

```{r}
# CLUSTERING WITH K-MEANS

# Clustering with 2 center (Fall vs. Spring Shoppers?)
#k3 <- kmeans(scaled_data[,c(-1)],3,nstart = 50,iter.max = 15) # keep number of iter.max=15 to ensure the algorithm converges and nstart=50 to #ensure that at least 50 random sets are choosen  
#k3

#fviz_cluster(k3, data=scaled_data[,c(-1)],ellipse.type = "convex")+ theme_minimal() # the clusters are hard to depict since it is in 2D. Use a confusion matrix decipher the k-means

```

```{r}
# cluster matrix
print(knitr::kable(table(scaled_data[,1],k2$cluster), format="markdown"))

# reattach assigned clusters to original data
cluster_IDs <- table(scaled_data[,1],k2$cluster) # store results of k-means in new dataframe
cluster_IDs <- as.data.frame(cluster_IDs)
colnames (cluster_IDs) <- c("Cust_ID", "Segment", "Frequency")
cluster_IDs <- cluster_IDs[!(cluster_IDs$Frequency==0),] # only use frequency when not equal to 0!

data_clusters <- merge(cust_infoDF, cluster_IDs[, c("Cust_ID", "Segment")], by="Cust_ID")
#write.csv(data_clusters, file = "2_segmentats_r.csv")

# create tables for each segment
segment1 <- data_clusters[!(data_clusters$Segment==2),]
segment2 <- data_clusters[!(data_clusters$Segment==1),]
```

```{r Segment Summaries}
segment1_salesdata <- segment1[,3:134]
is.na(segment1_salesdata) <- segment1_salesdata==0 #make 0's na's so they don't negatively skew the avg
colMeans(segment1_salesdata,  na.rm = TRUE, dims = 1)

segment2_salesdata <- segment2[,3:134]
is.na(segment2_salesdata) <- segment2_salesdata==0 #make 0's na's so they don't negatively skew the avg
colMeans(segment2_salesdata,  na.rm = TRUE, dims = 1)

segment1_demographics <- segment1[,135:170]
summary(segment1_demographics)

segment2_demographics <- segment2[,135:170]
summary(segment2_demographics)

# count occupation codes in each segment
library(plyr)
count(segment1_demographics, "OccupCd")
count(segment2_demographics, "OccupCd")
```

```{r Coding Customers by Channel}
retail_cust <- cust_infoDF[c("Cust_ID", "RetF07Trips", "RetS07Trips", "RetF06Trips", "RetS05Trips","RetF05Trips", "RetS05Trips","RetF04Trips", "RetS04Trips", "RetPre04Trips")]

retail_cust$retail=rowSums(retail_cust[,c(-1)])

retail_cust$retail[retail_cust$retail!= 0] <- 1
retail_cust$retail

internet_cust <-cust_infoDF[c("Cust_ID","IntF07Orders", "IntS07Orders", "IntF06Orders", "IntS06Orders","IntF05Orders", "IntS05Orders","IntF04Orders", "IntS04Orders", "IntPre04Orders")]

internet_cust$internet=rowSums(internet_cust[,c(-1)])

internet_cust$internet[internet_cust$internet!= 0] <- 1
internet_cust$internet

catalog_cust <-cust_infoDF[c("Cust_ID","CatF07Orders", "CatS07Orders", "CatF06Orders", "CatS06Orders","CatF05Orders", "CatS05Orders","CatF04Orders", "CatS04Orders", "CatPre04Orders")]

catalog_cust$catalog=rowSums(catalog_cust[,c(-1)])

catalog_cust$catalog[catalog_cust$catalog!= 0] <- 1
catalog_cust$catalog


# DF for First Channel to code columns
FirstChannel_DF <- cust_infoDF[c("Cust_ID", "FirstChannel")]

FirstChannel_DF$Retail <-FirstChannel_DF$FirstChannel
FirstChannel_DF$Retail[FirstChannel_DF$Retail!= 3] <- 0
FirstChannel_DF$Retail
FirstChannel_DF$Retail[FirstChannel_DF$Retail!= 0] <- 1
FirstChannel_DF$Retail

FirstChannel_DF$Internet<-FirstChannel_DF$FirstChannel
FirstChannel_DF$Internet[FirstChannel_DF$Internet!= 2] <- 0
FirstChannel_DF$Internet
FirstChannel_DF$Internet[FirstChannel_DF$Internet!= 0] <- 1
FirstChannel_DF$Internet

FirstChannel_DF$Catalog<-FirstChannel_DF$FirstChannel
FirstChannel_DF$Catalog[FirstChannel_DF$Catalog!= 1] <- 0
FirstChannel_DF$Catalog


ChannelLoyalty_DF <- cbind(cust_infoDF$Cust_ID, FirstChannel_DF$Retail, FirstChannel_DF$Internet, FirstChannel_DF$Catalog, cust_infoDF$FirstDollar, cust_infoDF$StoreDist, cust_infoDF$Email,cust_infoDF$AgeCode,cust_infoDF$IncCode,cluster_IDs$Segment,retail_cust$retail,internet_cust$internet,catalog_cust$catalog)

colnames(ChannelLoyalty_DF) <- c("Cust_ID", "FC_Retail", "FC_Internet", "FC_Catalog", "FirstDollar", "StoreDist", "Email", "Age", "Income", "Segment", "Retail", "Internet", "Catalog")

sum(catalog_cust$catalog)+sum(internet_cust$internet)+sum(retail_cust$retail)

#write.csv(ChannelLoyalty_DF, file = "channelloyalty.csv")
```


```{r}
Orders <- read.csv("DMEFExtractOrdersV01.CSV")
Orders_upduped<- Orders
Orders_upduped$OrderDate = substr(Orders_upduped$OrderDate ,1,nchar(Orders_upduped$OrderDate)-4) #make date only year
Orders_upduped<- Orders_upduped[Orders_upduped$OrderDate ==2007,] #keep only values = 2007
duplicated(Orders_upduped$Cust_ID)

Orders_upduped <- Orders_upduped[!duplicated(Orders_upduped$Cust_ID), ] #get rid of duplicates
duplicated(Orders_upduped$Cust_ID) #check

internet_orders <-cust_infoDF[c("Cust_ID","IntF07Orders", "IntS07Orders")]

internet_orders$orders=rowSums(internet_cust[,c(-1)])

internet2007 <- merge(internet_orders[,c("Cust_ID","orders")], Orders_upduped[, c("Cust_ID", "OrderDate")], by="Cust_ID")

write.csv(internet2007, file = "internet2007.csv")

```